Adaptive Sequential Model Selection

### 2015-05-08 paper.tex
commented out old sequential section
introduced separation and the sufficient filtration

\begin{example}[Principal Components Analysis]
  As a second motivating example, consider model selection for   principal components analysis. In that case we are given a data matrix $X \in \R^{n\times d}$, with which we form a sample covariance matrix
\[
S = \frac{1}{n-1} \sum_{i=1}^n(x_i - \bar x)^2
\]
The first $d$ principal component loadings are the first $d$ eigenvectors of $S$, which call $u_1,\ldots, u_d$. These induce a sequence of nested Wishart models:
\[
M_0 \sub M_1 \sub \cdots \sub M_d
\]
in which
\begin{equation}
  M_k:\; (n-1) S \sim W_d\left(\lambda_0 I_d + \sum_{\ell=1}^k     \lambda_\ell u_i u_i', \;\;\; n-1\right).
\end{equation}
This problem was studied in \citet{choi2014selecting}.
\end{example}

@@

To begin, we consider the problem of constructing a valid selective $p$-value for a particular step. At step $k$, we construct $p_k(Y)$ to test
\[
  H_{0,k}:\; F \in M_{k-1}(Y)
  \quad \text{ vs. } \quad
  H_{1,k}:\; F \in M_k(Y) \setminus M_{k-1}(Y).
\]
The main complication arises from the fact that the null and alternative hypotheses are random. To simplify matters, we first consider deriving $p$-values for a fixed candidate pair $m_{k-1} \sub m_k$.

The random variable  $p_{k,m_{k-1},m_k}(Y)$ is a valid {\em selective $p$-value} for the fixed candidate pair $(m_{k-1},m_k)$ if it is stochastically larger than uniform under the (fixed) null $m_{k-1}$, given that the pair $(m_{k-1},m_k)$ is selected. That is,
\[
\P_F\left(p_{k,m_{k-1},m_k}(Y) \leq \alpha \mid M_{k-1}(Y) = m_{k-1}\right) 
\leq \alpha, \quad \forall F\in m_{k-1}, \alpha \in [0,1].
\]
Given selective $p$-values for each fixed candidate pair, we can construct a combined $p$-value for the random null $(M_{k-1},M_k)$: 
\[
p_k(y) = p_{k, M_{k-1}(y), M_k(y)}(y),
\]
which is a valid $p$-value on the event $\{F \in M_{k-1}(Y)\}$:
\[
\P_F\left(p_k \leq \alpha \mid M_{k-1}, \;M_k, 
  \;F\in M_{k-1}\right) \leq \alpha, \quad \forall \alpha \in [0,1].
\]
Recall that, as usual, $F$ is not random, but $M_{k-1}$ is.

@@

\citet{gsell2013sequential} propose another stopping rule that controls the model-wise FWER, which they called {\em strong stop}. Define
\[
  \hk_{S}(Y) = \max\left\{k \in \{1,\ldots,d\} :\;
    \exp\left(\sum_{\ell=k}^d \frac{\log p_\ell}{\ell}\right) 
    \leq \frac{\alpha k}{d}\right\}
\]
Even if $p_k$ is a little larger than $\alpha$, say $\alpha=0.05$ and $p_k=0.06$, strong stop can still reject $M_{k-1}$ if the next three or four $p$-values are also relatively small.

\citet{gsell2013sequential} show that if the completion index $k_0$ is fixed, then $\hk_S$ controls the FWER at level $\alpha$, as long as the null $p$-values are independent given the non-null ones. Specifically, they require that
\begin{equation}\label{eq:indepCond_fixed_k0}
\P(p_{k_0+1} \leq \alpha_{k_0+1}, \ldots, p_d \leq \alpha_d
\mid p_1, \ldots, p_{k_0}) \leq \prod_{\ell=k_0+1}^d \alpha_\ell
\end{equation}

For random $k_0$, it is easy to see that the $p$-values simply need to satisfy~\eqref{eq:indepCond_fixed_k0} conditional on the value of $k_0$:
\begin{equation}\label{eq:indepCond_random_k0}
  \P(p_{k+1} \leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d
  \mid p_1, \ldots, p_{k_0}, \; k_0 = k) \leq \prod_{\ell=k+1}^d \alpha_\ell
\end{equation}
If~\eqref{eq:indepCond_random_k0} holds, then strong stop controls theprobability of $V>0$ conditionally on $k_0$, and thus also marginally.

@@

\citet{gsell2013sequential} propose another stopping rule, {\em forward stop}, that controls the model-wise FDR:
\[
  \hk_{F}(Y) = \max\left\{k \in \{1,\ldots,d\} :\;
    -\frac{1}{k}\sum_{\ell=1}^k \log(1-p_\ell) \leq \alpha\right\}
\]
If $p_k$ is uniform then  $-\log(1-p_k)$ is an exponential random variable. If all of the null $p$-values are uniform and the others are zero, then only the null ones contribute to the sum, and
\[
\widehat{\text{FDR}}_k = -\frac{1}{k}\sum_{\ell=1}^k \log(1-p_\ell)
\]
is a Gamma random variable with mean $V/k$ and variance $V/k^2$. We stop the last time the estimate of FDR is less than $\alpha$. Forward stop requires the same condition independence condition~\eqref{eq:indepCond_random_k0} as strong stop does.

\subsection{Saturated- and Selected-Model $p$-Values in Regression}

If we wish to use strong stop and forward stop, we need the $p$-values to satisfy~\eqref{eq:indepCond_random_k0}, but there is no guarantee that single-step selective $p$-values are independent of each other.

In fact, saturated-model $p$-values typically are not independent. Continuing with Example~\ref{ex:bivariate}, Table~\ref{tab:bv_twoWayTable} shows a two-way contingency table for the saturated-model $p$-values $(p_1(Y), p_2(Y))$, binned into cells of height and width 0.2. Because the global null is true, both $p$-values are uniform by construction, but the $p$-values are strongly dependent, with correlation $-48\%$.

@@

By contrast, the selected-model $p$-values $(p_1,p_2)$ are independent. To see why, recall that, on the selection event $A$, $p_1$ is only a function of $Y_1$. Moreover, $p_2$ is a selective $p$-value for comparing the null
\[
M_1:\; Y \sim \cN\left(\binom{\mu_1}{0}, I_2\right)
\]
against the alternative
\[
M_2:\; Y \sim \cN\left(\mu, I_2\right), \text{ with } \mu_2 \neq 0.
\]
Because $\mu_1$ is a nuisance parameter in $M_1$, an exact selected-model test must condition on the corresponding sufficient statistic $Y_1$. But that means $p_2$ is uniform given $Y_1$, and therefore also uniform given $p_1(Y_1)$.

A generalization of this argument shows that the $p$-values for the saturated-model tests are always independent:

\begin{proposition}
  The selected-model equal-tailed or UMPU $p$-values after forward-stepwise linear regression. satisfy~\eqref{eq:indepCond_random_k0}.
\end{proposition}
This is a simple consequence of the much more general results proven in Section~\ref{sec:sufficiency}, but it is instructive to see the proof in the relatively concrete case of linear regression.
\begin{proof}
  At step $k$, we are testing the null hypothesis that $\beta_j=0$ in     the model $Y\sim \cN(X_E\beta, I_2)$, for some active set $E\sub [p]$ of size $k$, and some variable $j\in E$. Let $A_{j,E}$ denote the selection event for a particular $j$ and $E$, and let $m_{k-1}(j,E)$ denote the null model with active set $X_{E\setminus j}$.

On $A_{j, E}$, the selected-model $p$-value $p_k$ is based on the null distribution of the test statistic $X_j'Y$ conditional on the null sufficient statistics $X_{E\setminus j}'Y$ and on $Y \in A_{j,E}$.

Moreover, on $A_{j,E}$, the entire sequence of previous $p$-values $p_{[k-1]}$ depend only on the sufficient statistics $X_{E\setminus j}'Y$ of $m_{k-1}(j,E)$. To see why, \WFcomment{complete this proof.}

We have shown that, if $m_{k-1}(j,E)$ is true, then $p_k$ is uniform and independent of $p_{[k-1]}$ on $A_{j,E}$ for each $j$ and $E$. That is,
\[
\P_F\left(p_{k} \leq \alpha \mid p_{[k-1]}, Y \in A_{j,E} \right) = \alpha \quad \forall F \in m_{k-1}.
\]
It follows that 
\[
\P_F\left(p_{k} \leq \alpha \mid p_{[k-1]}, F \in M_{k-1} \right) = \alpha \quad \forall F.
\]
\end{proof}
The proof turns on two properties of selected-model $p$-values, and one property of forward stepwise regression.

As for the $p$-values, we used the fact that $p_{k,m_{k-1},m_k}$ is uniform conditional on the sufficient statistics of the null model $m_{k-1}$. This is true for selected-model and saturated-model $p$-values alike. We also used the fact that $p_{k,m_{k-1},m_k}$ follows the {\em sufficiency principle} --- that is, it depends only on the sufficient statistics of $m_k$. This is {\em not} true of the saturated-model $p$-values: to compute them, we must know $\proj_{\eta}^\perp Y$, which depends on statistics that are ancillary for $m_{k-1}$.

We also used the fact that, in the forward stepwise algorithm, if $m_k$ is the $k$th model then, on the event $M_k=m_k$, all the models up to step $k$ depend only on sufficient statistics of $m_k$. This is a sort of {\em sub-path sufficiency principle} which many other algorithms satisfy, as we will see in Section~\ref{sec:modelSPSP}.

\subsection{Conditions for Independent $p$-Values}\label{sec:sufficiency}

To obtain a sequence of $p$-values, we must specify what model selection algorithm $M_{[d]}$ to use, as well as how to construct $p$-values at each step. In this section we consider what conditions are sufficient to guarantee that the resulting $p$-value sequence is independent. Sections~\ref{sec:pvalSP}--\ref{sec:modelSPSP} discuss sufficient conditions on the $p$-values and the selection algorithm under which single-step $p$-values are automatically independent. Section~\ref{sec:selectionVariables} discusses how we can create independent $p$-values by conditioning on finer selection variables at each step.

\subsubsection{Generalized Selected-Model $p$-Values}\label{sec:pvalSP}

For a candidate pair $(m_{k-1},m_k)$ of models to test at step $k$, define the selection event $A = \{M_{k-1},M_k\}$. We say $p(Y)$ is a {\em generalized selected-model $p$-value} for testing $m_{k-1}$ against $m_k$ at step $k$ if it is ``based on'' the law
\[
\L_{m_{k-1}}\left(T_k \;\;\mid\;\; T_{k-1},\; M_{k-1}=m_{k-1},\;M_k=m_k\right)
\]
for some minimal sufficient statistics $T_{k-1}(Y)$ and $T_k(Y)$ of $m_{k-1}$ and $m_k$, respectively. More precisely, $p(Y)$ must be uniform or conservative conditional on $T_{k-1}(Y)$, and it must be a function of $T_k(Y)$ on $A$.

\subsubsection{A Sub-Path Sufficiency Principle for Model Selection Procedures}\label{sec:modelSPSP}

We say that a selection algorithm $M_{[d]}(\cdot)$ satisfies the {\em sub-path sufficiency principle} (henceforth SPSP) if, on the event $M_k(Y)=m_k$, the sequence of models $M_{[k]}$ depends only on sufficient statistics of $m_k$. Again, if

@@

For $y\in A$ and $\ell \leq k$, we have $E_\ell(y) \sub E$. This means that the maximizer in \eqref{eq:forwardDef_start} is attained for some $j\in E$. So, for $\ell \leq k$, on $A$,
\begin{align}
  j_\ell &= \argmax_{j\in E} \;\;\sup \left\{L(\theta; Y):\;
    \theta\in\Theta(E_{\ell-1} \cup \{j\})\right\} \\
  &= \argmax_{j\in E} \;\;\sup \left\{L_T(\theta; T(Y)):\;
    \theta\in\Theta(E_{\ell-1} \cup \{j\})\right\}. \label{eq:onlyT}
\end{align}
Replacing $L$ with $L_T$ in~\eqref{eq:onlyT} is justified because $\Theta(E_{\ell - 1} \cup \{j\}) \sub \Theta(E)$ when $E_{\ell-1}\cup E$ and $j\in E$.

@@

\begin{comment}
\WFcomment{There is a filtration interpretation when you have the appropriate sufficiency properties.} Let $\sF_{k,\ell}$ denote the $\sigma$-algebra generated by $M_{[k]}$ and $p_{[\ell]}$.
\begin{align*}
  \sF_{k,\ell} &= \sF(M_{[k]},p_{[\ell]})\\
  \sF_0 &\underlabel_{\text{selection } 1} \sF_{1,0} \underlabel_{\text{inference } 1}
  \sF_{1,1} \;\;\sub \cdots \sub\;\;
  \sF_{d-1,d-1} \underlabel_{\text{selection } d} \sF_{d,d-1}
  \underlabel_{\text{inference } d} \sF_{d,d}
\end{align*}
\end{comment}

@@

We say that $p_{k,m_{k-1},m_k}$ is a selective $p$-value with respect to selection variable $S(Y)$ if it is uniform under $F\in m_k$ given $S$.

To achieve joint error control properties among different tests, we may need to condition on a finer variable.

@@

>>>>

\begin{example}[Principal Components Analysis]
  As a second motivating example, consider model selection for   principal components analysis. In that case we are given a data matrix $X \in \R^{n\times d}$, with which we form a sample covariance matrix
\[
S = \frac{1}{n-1} \sum_{i=1}^n(x_i - \bar x)^2
\]
The first $d$ principal component loadings are the first $d$ eigenvectors of $S$, which call $u_1,\ldots, u_d$. These induce a sequence of nested Wishart models:
\[
M_0 \sub M_1 \sub \cdots \sub M_d
\]
in which
\begin{equation}
  M_k:\; (n-1) S \sim W_d\left(\lambda_0 I_d + \sum_{i=1}^k     \lambda_i u_i u_i', \;\;\; n-1\right).
\end{equation}
This problem was studied in \citet{choi2014selecting}.
\end{example}

@@

To begin, we consider the problem of constructing a valid selective $p$-value for a particular step. At step $k$, we construct $p_k(Y)$ to test
\[
  H_{0,k}:\; F \in M_{k-1}(Y)
  \quad \text{ vs. } \quad
  H_{1,k}:\; F \in M_k(Y) \setminus M_{k-1}(Y).
\]
The main complication arises from the fact that the null and alternative hypotheses are random. To simplify matters, we first consider deriving $p$-values for a fixed candidate pair $m_{k-1} \sub m_{k-1}$.

The random variable  $p_{k,M}(Y)$ is a valid {\em selective $p$-value} for the fixed candidate pair $(m_{k-1},m_k)$ if it is stochastically larger than uniform under the (fixed) null $m_{k-1}$, given that the pair $(m_{k-1},m_k)$ is selected. That is,
\[
\P_F\left(p_{k,M}(Y) \leq \alpha \mid M_{k-1}(Y) = m_{k-1}\right) 
\leq \alpha, \quad \forall F\in m_{k-1}, \alpha \in [0,1].
\]
Given selective $p$-values for each fixed candidate pair, we can construct a combined $p$-value for the random null $(M_{k-1},M_k)$: 
\[
p_k(y) = p_{k, M_{k-1}(y), M_k(y)}(y),
\]
which is a valid $p$-value on the event $\{F \in M_{k-1}(Y)\}$:
\[
\P_F\left(p_k \leq \alpha \mid M_{k-1}, \;M_k, 
  \;F\in M_{k-1}\right) \leq \alpha, \quad \forall \alpha \in [0,1].
\]
Recall that, as usual, $F$ is not random, but $M_{k-1}$ is.

@@

\citet{gsell2013sequential} propose another stopping rule that controls the model-wise FWER, which they called {\em strong stop}. Define
\[
  \hk_{S}(Y) = \max\left\{k \in \{1,\ldots,d\} :\;
    \exp\left(\sum_{i=k}^d \frac{\log p_i}{i}\right) 
    \leq \frac{\alpha k}{d}\right\}
\]
Even if $p_k$ is a little larger than $\alpha$, say $\alpha=0.05$ and $p_k=0.06$, strong stop can still reject $M_{k-1}$ if the next three or four $p$-values are also relatively small.

\citet{gsell2013sequential} show that if the completion index $k_0$ is fixed, then $\hk_S$ controls the FWER at level $\alpha$, as long as the null $p$-values are independent given the non-null ones. Specifically, they require that
\begin{equation}\label{eq:indepCond_fixed_k0}
\P(p_{k_0+1} \leq \alpha_{k_0+1}, \ldots, p_d \leq \alpha_d
\mid p_1, \ldots, p_{k_0}) \leq \prod_{i=k_0+1}^d \alpha_i
\end{equation}

For random $k_0$, it is easy to see that the $p$-values simply need to satisfy~\eqref{eq:indepCond_fixed_k0} conditional on the value of $k_0$:
\begin{equation}\label{eq:indepCond_random_k0}
  \P(p_{k+1} \leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d
  \mid p_1, \ldots, p_{k_0}, \; k_0 = k) \leq \prod_{i=k+1}^d \alpha_i
\end{equation}
If~\eqref{eq:indepCond_random_k0} holds, then strong stop controls theprobability of $V>0$ conditionally on $k_0$, and thus also marginally.

@@

\citet{gsell2013sequential} propose another stopping rule, {\em forward stop}, that controls the model-wise FDR:
\[
  \hk_{F}(Y) = \max\left\{k \in \{1,\ldots,d\} :\;
    -\frac{1}{k}\sum_{i=1}^k \log(1-p_i) \leq \alpha\right\}
\]
If $p_k$ is uniform then  $-\log(1-p_k)$ is an exponential random variable. If all of the null $p$-values are uniform and the others are zero, then only the null ones contribute to the sum, and
\[
\widehat{\text{FDR}}_k = -\frac{1}{k}\sum_{i=1}^k \log(1-p_i)
\]
is a Gamma random variable with mean $V/k$ and variance $V/k^2$. We stop the last time the estimate of FDR is less than $\alpha$. Forward stop requires the same condition independence condition~\eqref{eq:indepCond_random_k0} as strong stop does.

\subsection{$p$-Values and Independence}

Recall that $p_k$ is a valid selective $p$-value for $M_{k-1}\sub M_\infty$ if, for all $F\in M_\infty$,
\[
\P_F(p_k \leq \alpha \mid M_{k-1},\; F\in M_{k-1}) \leqAS \alpha.
\]
Valid single-step selective $p$-values do not generally result in $p$-values that are independent from one step to the next.

For example, saturated-model $p$-values are generically non-independent. Continuing with Example~\ref{ex:bivariate}, Table~\ref{tab:bv_twoWayTable} shows a two-way contingency table for the saturated-model $p$-values $(p_1(Y), p_2(Y))$, binned into cells of height and width 0.2, simulated under the global null $\mu=0$. Because $k_0=0$, both $p$-values are uniform by construction, but the $p$-values are strongly dependent, with correlation $-48\%$. By contrast, we will see that the selected-model $p$-values $(p_1,p_2)$ are independent under the global null, a consequence of Proposition~\ref{prop:selectiveModel}.

@@

First, note that $p_2$ is a selective $p$-value for comparing the null with only variable 1,
\[
M_1:\; Y \sim \cN\left(\binom{\mu_1}{0}, I_2\right),
\]
against the alternative with variables 1 and 2:
\[
M_2:\; Y \sim \cN\left(\mu, I_2\right), \text{ with } 
\mu_2 \neq 0.
\]
To eliminate the nuisance parameter $\mu_1$, the test conditions on $Y_1$. Thus, $p_2$ is uniform given $Y_1$ on $A$. Second, note that on $A$, $p_1$ is a function only of $Y_1$. Thus, $(p_1,p_2)$ are independent uniform variables given that variable 1 is chosen first. By a similar argument, they would also be independent uniforms if $Y_2$ were chosen first; thus, $(p_1, p_2)$ are marginally uniform under the global null.

We now introduce a simple sufficient condition to verify that a given procedure yields independent $p$-values. We say that a filtration $\sF_{[d]}$ with $M_{[k]} \in \sF_k$ {\em separates} the $p$-values $p_{[d]}$ if 
\begin{enumerate}
\item $p_k(Y)$ is conservative given $\sF_{k-1}$ 
  when $F\in M_{k-1}$, and
\item $p_k(Y)$ is measurable with respect to $\sF_k$.
\end{enumerate}
If we think of $\sF_k$ representing information available at step $k$, then the first condition means that $p_k$ excludes whatever evidence may have accrued against the null by step $k-1$, and the second means that information revealed after step $k$ plays no role in determining $p_k$. In that sense, $\sF_k$ forms a ``wall of separation'' between the $p$-values up to $k$ and the ones after $k$. As a result, separated null $p$-values are independent.

\begin{proposition}[Independence of 
  Separated $p$-Values]\label{prop:jointConserv}

  Let $p_{[d]}$ be selective $p$-values for $M_{[d]}$, 
  separated by $\sF_{[d]}$.

  If the $p$-values are exact then $p_{k+1}, \ldots, p_d$ are
  uniform given $\sF_{k_0}$ on the event $\{k_0=k\}$.

  If they $p$-values are conservative, then for
  $\alpha_{k+1},\ldots,\alpha_d \in [0,1]$,
  \begin{equation}\label{eq:jointConserv}
  \P_F\left(p_{k+1}\leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d\;
    \mid\; k_0 \leq k, \; \sF_k\right) \leqAS \prod_{i=k+1}^d
  \alpha_i.
  \end{equation}
\end{proposition}
Equation~\ref{eq:jointConserv} is slightly weaker than independence, but it guarantees that both Strong Stop and Forward Stop control their target error rates.

\begin{proof}
  We first prove~\eqref{eq:jointConserv}
  by induction. Define $B_i = \{p_i \leq \alpha_i\}$. 
  The base case is
  \[
  \P_F\left(B_d \mid \sF_{d-1}\right)1_{\{k_0 \leq d-1\}} \leqAS \alpha_d,
  \]
  which is true by construction of $p_d$. 
  For the inductive case, note that the
  conditional probability on the left-hand side
  of~\eqref{eq:jointConserv} is
  \begin{align*}
    \P_F\left(B_{k+1}, \ldots, B_d
      \mid \sF_k\right)1_{\{k_0 \leq k\}} 
    &\eqAS \E_F\bigg[ 1_{B_{k+1}} 
    \P_F\left(B_{k+2}, \ldots, B_d
      \mid \sF_{k+1}\right)1_{\{k_0 \leq k+1\}}
    \mid \sF_k\bigg]1_{\{k_0 \leq k\}}\\
    &\leqAS \P_F\left[ B_{k+1}
      \mid \sF_k\right]1_{\{k_0 \leq k+1\}}\prod_{i=k+2}^d \alpha_i
    \quad\;\;\leqAS\;\; \prod_{i=k+1}^d \alpha_i
  \end{align*}

  If the $p_k$ are exact then the above 
  inequalities become equalities, implying uniformity.
\end{proof}

\subsection{Sufficient Filtration}

Assume that each candidate model has a minimal sufficient statistic $T(y; M)$, and write
\[
T_k(Y) = T(Y; M_k(Y)).
\]
For example, in linear regression with known $\sigma^2$, the complete sufficient statistic for model $M(E)$ is $X_E'Y$, and the sufficient statistic at step $k$ is $T_k = X_{E_k}'Y$.

We will concern ourselves mainly with the {\em sufficient filtration} for path $M_{[k]}$, defined as
\[
\sF_k = \sF(M_{[k]}, T_k),
\]
The sufficient filtration separates $p_{[d]}$ if $p_k$ is conservative given $\sF_{k-1}$ and $\sF_k$-measurable.\footnote{For readers unfamiliar with $\sigma$-algebra notation, this statement can be read as ``$p_k$ is conservative given $(M_{[k-1]}, T_{k-1})$ and is a function of $(M_{[k]}, T_k)$''}

To be a valid selective $p$-value for a single step, $p_k$ only had to condition on $M_{k-1}$. Thus, it is a little more stringent to additionally require $p_k$ to condition on the entire subpath $M_{[k-1]}$ as well as $T_{k-1}$.

In most cases of interest, however, the two requirements are equivalent. First, we typically construct $p_k$ by conditioning on $T_{k-1}$. In fact, $T_{k-1}$ is complete sufficient, then any exact test must condition on $T_{k-1}$ \WFcomment{should we bother proving this?}. Second, most path algorithms of interest (including forward stepwise regression, the ever-active LASSO path, and their generalizations) satisfy a {\em subpath sufficiency principle} (henceforth, SSP): that once we know $M_k$ and its sufficient statistics, we can reconstruct the model path up to step $k$. Thus, most single-step $p$-values are already conservative conditional on $\sF_{k-1}$.

The requirement that $p_k$ be $\sF_k$-measurable has more bite. For example, it excludes saturated-model tests in linear regression. Computing the cutoff for the saturated-model test requires us to know $\proj_{\eta}^\perp Y$, which is not $\sF_k$-measurable.

By contrast, selected-model tests of $M_{k-1}$ against $M_k$ are always $\sF_k$-measurable, because they are based on the law
\begin{equation}\label{eq:selModel}
\L\left(X_{j_k}'Y \mid X_{E_k-1}'Y, E_{k-1}, j_k\right),
\end{equation}
and all of the random variables appearing in~\eqref{eq:selModel} are $\sF_k$-measurable.

\begin{comment}

\begin{theorem}
  Assume that $T(Y; M)$ is a complete sufficient statistic, 
  that $p_k$ are exact selective $p$-values, and that 
  $M_{[d]}$ satisfies EPIC.

  Then, the sufficient filtration separates $p_{[d]}$ if and only if 
  $p_k \in \sF_k$.
\end{theorem}

@@

For $y\in A$ and $i \leq k$, we have $E_i(y) \sub E$. This means that the maximizer in \eqref{eq:forwardDef_start} is attained for some $j\in E$. So, for $i \leq k$, on $A$,
\begin{align}
  j_i &= \argmax_{j\in E} \;\;\sup \left\{L(\theta; Y):\;
    \theta\in\Theta(E_{i-1} \cup \{j\})\right\} \\
  &= \argmax_{j\in E} \;\;\sup \left\{L_T(\theta; T(Y)):\;
    \theta\in\Theta(E_{i-1} \cup \{j\})\right\}. \label{eq:onlyT}
\end{align}
Replacing $L$ with $L_T$ in~\eqref{eq:onlyT} is justified because $\Theta(E_{i - 1} \cup \{j\}) \sub \Theta(E)$ when $E_{i-1}\cup E$ and $j\in E$.

@@

@@

We say that $p_{k,M}$ is a selective $p$-value with respect to selection variable $S(Y)$ if it is uniform under $F\in m_k$ given $S$.

To achieve joint error control properties among different tests, we may need to condition on a finer variable.
\end{comment}

@@

\begin{comment}
Sections~\ref{sec:pvalSP}--\ref{sec:modelSPSP} discuss sufficient conditions on the $p$-values and the selection algorithm under which single-step $p$-values are automatically independent. Section~\ref{sec:selectionVariables} discusses how we can create independent $p$-values by conditioning on finer selection variables at each step.

\end{comment}

\begin{comment}
\WFcomment{There is a filtration interpretation when you have the appropriate sufficiency properties.} Let $\sF_{k,i}$ denote the $\sigma$-algebra generated by $M_{[k]}$ and $p_{[i]}$.
\begin{align*}
  \sF_{k,i} &= \sF(M_{[k]},p_{[i]})\\
  \sF_0 &\underlabel_{\text{selection } 1} \sF_{1,0} \underlabel_{\text{inference } 1}
  \sF_{1,1} \;\;\sub \cdots \sub\;\;
  \sF_{d-1,d-1} \underlabel_{\text{selection } d} \sF_{d,d-1}
  \underlabel_{\text{inference } d} \sF_{d,d}
\end{align*}
\end{comment}
